---
title: "Report on Examing the Effect of Income, Education on Domestic Incidence Rate"
author: Junyou Chen
date: "2/23/2022"
output: 
  github_document:
    toc: true
    
---

```{r setup, include=FALSE}
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, out.width="120%")
if(!require(GGally)) install.packages("GGally", repos = "https://github.com/ggobi/ggally")

## install packages and load libraries
install.packages("gridExtra")
library(here)
library(tidyverse)
library(readxl)
library(knitr)
library(gridExtra) 
library(knitr)
library(GGally)
library(broom)

theme_set(theme_minimal())

```

# 1. Introduction

## 1.1 Research question
'Family incidents' recorded by Victoria Police increased by 6.7 percent from 82,651 in 2018–19 to 88,214 in 2019–20 (a 5% increase in the rate of incidents per 100,000 people), affecting tens of thousands of families across Victoria. As a result, I'd like to use the datasets I've acquired to investigate possible causes of domestic violence across Victoria, allowing policymakers to make more informed decisions.Therefore, my research question is, Do income level and education level have effects on domestic incidence rate? To examine the possible impact of income and education on domestic incidence, I have chosen Victoria state, Australia, as my research setting to conduct the analysis. I wish to find possible correlations between income, education and family violence for policy makers to make more effective policies targeting the issue.

# 2. Data

## 2.1 Dataset
As the census data for year 2021 is not yet available in [Aurin](https://portal.aurin.org.au/), I've chosen most recent census data in 2016 to conduct the analysis. To keep the consistency of the data, I also ajusted the domestic violence data to year 2016.

**V_Police_Data_2012-17.xlsx**
Bureau of crime statistics data sheet on victims of domestic violence recorded by [Victorian police](https://www.crimestatistics.vic.gov.au/family-violence-data-portal/download-data-tables).This includes data recorded between July 2012 and July 2017, and 21 data sheets on many different aspects of domestic violence cases reported to the police, including the gender and age of affected members, children's participation in domestic violence cases, etc. However, the main data of this study are **Table 2** and  **Table 3**.

**Table 2** has been stored as family_incidence, including family event data recorded in police areas and local government areas from July 2012 to June 2017.

**Table 3** has been stored as family_incidence_rate, including the data of household accident rate per 100000 population divided by police district and local government area from July 2012 to June 2017.

**AURIN_dataset.csv**
A datasheet from the Australian bureau of statics (ABS) that records the income, education, employability and health statistics by LGA (local government area). For this investigation, we are only interested in certain variables including (for a person over the age of 15): completed school years and the weekly total personal income for residence in victoria, Australia. The data are collected by the ABS from the 2016 census of Population and Housing.
 
```{r `clean data`,include=FALSE}
#import the data
family_incidence_raw <- read_excel(here("V_Police_Data_2012-17.xlsx"), 
                               sheet = 3, skip = 13,range = "C14:H96")
family_incidence_rate_raw <- read_excel(here("V_Police_Data_2012-17.xlsx"), 
                                   sheet = 4, skip = 13,range = "C14:H96")
income_education_raw <- read_csv(here("AURIN_dataset_true.csv"))

```

```{r `table for number of family incidence in each region of victoria`,include=FALSE}
#create a table for domestic incidence number happened in each region of victory in year2016
family_incidence <- family_incidence_raw%>%
  mutate("region_name" = `Local government area`) %>%
  select("2015-16","region_name") %>%
  filter(region_name != "Total") %>%
  rename("family_incidents" = "2015-16") %>%
  data.frame() 
## to set region_name as factor
family_incidence$region_name <- as.factor(family_incidence$region_name)

family_incidence
```

```{r `table for the rate of family incidence in each region of victoria`,include=FALSE}
#create a table for domestic incidence per 10,000 persons happened in each region of victory in year2016
family_incidence_rate <- family_incidence_rate_raw %>%
  mutate("region_name" = `Local government area`) %>%
  select("2015-16","region_name") %>%
  filter(region_name != "Total") %>%
  rename("fincident_rate" = "2015-16") %>%
  data.frame()
## to set region_name as factor
family_incidence_rate$region_name <- as.factor(family_incidence_rate$region_name)

family_incidence_rate
```
```{r `table summarizing income level and high school graduation rate`,include=FALSE}
###create a table summarizing income levels and high_school graduation rate in each region of victory in year2016
wkly_income <- c("inc_over15_wkly_negative_%", "inc_over15_wkly_zero_%",
         "inc_over15_wkly_1_499_%","inc_over15_wkly_500_999_%",
         "inc_over15_wkly_1000_1999_%","inc_over15_wkly_2000_2999_%",
         "inc_over15_wkly_over_3k_%")
edu_year <- c("hg_edu_no_sch_%","hg_edu_comp_yr_8_or_blw_%","hg_edu_comp_yr_9_%",
         "hg_edu_comp_yr_10_%","hg_edu_comp_yr_11_%","hg_edu_comp_yr_12_%")

income_education <- income_education_raw %>%
  select("lga_name16", 
         starts_with("inc_over15"), 
         starts_with("hg")) %>%
  rename(region_name_pre = lga_name16) %>%
##to convert income and school year into less complicated categories. 
  mutate(low_income = 
        `inc_over15_wkly_negative_%`+`inc_over15_wkly_zero_%`+
        `inc_over15_wkly_1_499_%`+`inc_over15_wkly_500_999_%`,
         median_income = 
          `inc_over15_wkly_1000_1999_%` + 
          `inc_over15_wkly_2000_2999_%`,
        high_income = `inc_over15_wkly_over_3k_%`,
        graduated = `hg_edu_comp_yr_12_%`,
        ungraduated = 
        `hg_edu_no_sch_%` + `hg_edu_comp_yr_8_or_blw_%` +
          `hg_edu_comp_yr_9_%` +`hg_edu_comp_yr_10_%` +
          `hg_edu_comp_yr_11_%`) %>%
  select(region_name_pre, low_income, median_income, high_income, 
         graduated, ungraduated)
  
## to delete last 4 letters for column `region_name_pre`
income_education$region_name_pre <- substr(income_education$region_name_pre, 
                                           start = 1, 
                                           stop = nchar(income_education$region_name_pre) - 4)
## rename column
income_education <- income_education %>%
  rename(region_name = region_name_pre)
income_education
  
```
## 2.2 Data Wrangling
There are in total 79 Local Government Area(LGA) in Victoria. The following head of the table joined the table from V_Police_Data_2012-17.xlsx and AURIN_dataset.csv, and summarized different income levels and high school graudation rate in each region of Victoria. People earning weekly negative to 999AUD are identified as low income family, people earning weekly 1000AUD to 2999AUD are identified as median income family while people earning weekly over 3000AUD are identified as high income family.
```{r}
## to combine all three dataframes
semi_data <- inner_join(income_education, family_incidence, by = "region_name")
full_data <- data.frame(inner_join(semi_data, family_incidence_rate, by = "region_name"))
kable(head(full_data))

```

# 3 Analysis

## 3.1 Descriptive Analysis

### 3.1.1 Domestic Incidence in Victoria by Local Government Area

As we can see from the graph, among 79 LGAs, Casey(4387) has the highest domestic incidence number in year 2016 while Queenscliffe(6) has the lowest domestic incidence number. The average number of domestic incidence happened in Victoria in 2016 is 985. 
```{r Domestic_Incidence_in_Victoria_by_LGA}
## drawing bars to layout Domestic Incidence in Victoria by Local Government Area
family_incidence %>%
  arrange(family_incidents) %>%
  ggplot(mapping = aes(reorder(region_name, family_incidents),
                       y = family_incidents), cex.axis = 0.8) +
  geom_col(fill = "steelblue") +
  labs(size = 0.8, title = "Figure 3.1.1: Domestic Incidence Number in Victoria by Local Government Area",
       subtitle = "(2015~2016)",
       x = "Area Name", 
       y = "Domestic Incidence Number",
       caption = "Source: Victorian Police") +
   ##change word size to make it more clear
  theme_gray(base_size = 6) +
  coord_flip()

```

### 3.1.2 Domestic Incidence per 100,000 in Victoria by Local Government Area

As we can see from the graph, among 79 LGAs, Latrobe(3681) has the highest number of domestic incidence per 100,000 persons in year 2016 while Queenscliffe(195.5) has the lowest. The average number of domestic incidence per 100,000 persosn happened in Victoria in 2016 is 1431. 
```{r Domestic_Incidence_per_100,000_in_Victoria_by_LGA}
## drawing bars to layout Domestic Incidence rate in Victoria by Local Government Area
family_incidence_rate %>%
  arrange(fincident_rate) %>%
  ggplot(mapping = aes(reorder(region_name, fincident_rate),
                       y = fincident_rate)) +
  geom_col(fill = "steelblue") +
  labs(title = "Figure 3.1.2: Domestic Incidence Rate per 100,000 in Victoria by Local Government Area",
       subtitle = "(2015~2016)",
       x = "Area Name", 
       y = "Domestic Incidence Rate per 100,000",
       caption = "Source: Victorian Police") +
  theme_gray(base_size = 6) +
  coord_flip()

```

### 3.1.3 Income Level in Victoria by Local Government Area

*(The columns might not add up to 100% due to non-response rate or have been tweaked done by the Australian Bureau of Statistics out of the concern to protect confidentiality)*

For income level in State of Victoria, among 79 LGAs, people of low-income occupied a large proportion in regards of total population. The average percent of low-income population for 79 LGAs in Victoria State is 63.6%, while 24.2% for median-income population and 2.5% for high-income population. 
```{r Income_Level_in_Victoria_by_LGA}
## draw bar graphs to show different income levels in different area governments of victoria. 
full_data %>%
  select(low_income,
         median_income, 
         high_income, 
         region_name) %>%
  ## to make it look better on the legend
  rename(`High Income` = high_income, 
         `Low Income` = low_income, 
         `median Income` = median_income) %>%
  pivot_longer(cols = 1:3, 
               names_to = "income", 
               values_to = "pct_income") %>%
  ggplot(mapping = aes(x = reorder(region_name,desc(pct_income)), 
                       y = pct_income, 
                       fill = income, order = income)) +
  geom_bar(stat = "identity", alpha = 0.9) +
  labs(title = "Figure 3.1.3: Income Level in Victoria by Local Government Area",
       subtitle = "(2016)",
       x = "Area Name", 
       y = "Percentage of Income Level",
       caption = "Source: Australian bureau of statics (ABS)") +
##to change the color to make it prettier 
  scale_fill_brewer(palette = "Set2") +
  theme_gray(base_size = 6) +
  coord_flip()

```

### 3.1.4 Highschool Graduation Rate by Local Government Area

*(The columns might not add up to 100% due to non-response rate or have been tweaked done by the Australian Bureau of Statistics out of the concern to protect confidentiality)*

For highschool graduation rate, in the graph blue represents percentage of highschool graduation rate, among 79 LGAs, Melbourne(77%) has the highest high school graduation rate while Gannawarra(26.1%) has the lowest. The average high school graduation rate for the state of Victoria is 49.9%.
```{r Highschool_Graduation_Rate_by_LGA}
## draw overlapsed bar graphs in order to show different income levels in different area governments of victoria. 
full_data %>%
  select(graduated, 
         ungraduated, 
         region_name) %>%
  ggplot(mapping = aes(x = reorder(region_name, graduated), y = graduated)) +
  geom_col(aes(y = graduated), fill = "steelblue", alpha = 0.5) +
  scale_fill_manual(name="Bar", values=cols) +
  labs(title = "Figure 3.1.4: Education Level in Victoria by Local Government Area",
       subtitle = "(2016)",
       x = "Area Name",
       y = "Percentage of Highschool Graduation Rate",
       caption = "Source: Australian bureau of statics (ABS)",
       colour = "graduated") +
  theme_gray(base_size = 6) +
  coord_flip() 

```


## 3.2 Correlation Analysis

### 3.2.1 Income level vs. Number of Domestic Incidence
Scatterplot can be plotted to examine the relationship between income levels and domestic incidence numbers per 100,000 people. The first graph examined the relationship between the percentage of population of low-income and domestic incidence numbers per 100,000 people. As the graph shows, generally speaking, higher rate of low-income population is correlated to more numbers of domestic incidence per 100,000 people. In other words, domestic incidence is more likely to happen in an area with higher percentage of low-income populations. Same scatterplot can be also plotted to examine the relationship between the percentage of population of high-income and domestic incidence numbers per 100,000 people to further verify the result from the opposite direction. As the second graph demonstrates, domestic incidence is less likely to happen in an area with higher percentage of high-income populations, which keeps in accordance with the initial graph. 
```{r Income_level_vs._Number_of_Domestic_Incidence}
plot1 <- full_data %>%
  select(low_income, fincident_rate) %>%
  ggplot(mapping = aes(x = low_income, y = fincident_rate)) +
  geom_point(size = 2, color = "steelblue", alpha = 0.7) +
  geom_smooth(color = "darkred") +
    labs(title = "Figure 3.2.1: Low-income Percentage vs. Domestic Incidence/100,000",
         subtitle = "(2016)",
       x = "Percentage of Population of Low-income", 
       y = "Domestic Incidence cases/100,000",
       caption = "Source: Australian bureau of statics (ABS), Victorian Police") +
  ##change title size to make it more clear
  theme(plot.title = element_text(size=6)) +
  theme_gray(base_size = 6)

plot2 <- full_data %>%
  select(high_income, fincident_rate) %>%
  ggplot(mapping = aes(x = high_income, y = fincident_rate)) +
  geom_point(size = 2, color = "steelblue", alpha = 0.7) +
  geom_smooth(color = "darkred") +
    labs(title = "High-income Percentage vs. Domestic Incidence/100,000",
         subtitle = "(2016)",
       x = "Percentage of Population of Low-income", 
       y = "Domestic Incidence cases/100,000",
       caption = "Source: Australian bureau of statics (ABS), Victorian Police") +
  theme(plot.title = element_text(size=6)) +
  theme_gray(base_size = 6)

grid.arrange(plot1, plot2, ncol = 2)
```
**Pearson's r**

To further examine the correlation between the income level and domestic incidence numbers per 100,000 people from statistical point, Pearson's r can be conducted. As the result shows, the correlation between the percentage of population of low-income and domestic incidence numbers per 100,000 people is statistically significant *(p<0.01)* and is considered moderate in terms of strength (r = 0.48). In other words, the percentage of population of low-income in the region is positively related to the number of domestic incidence per 100,000 persons. 
```{r}
##to conduct pearson's r to examine the relationship between lowincome and rate of domestic incidence and that of high school graduation rate and rate of domestic incidence. 
l <- list(
  c1 = cor.test( ~ low_income + fincident_rate, full_data),
  c2 = cor.test( ~ graduated + fincident_rate, full_data)
)
c_df <- map_dfr(l,tidy,.id = 'id')
c_df %>%
  filter(id == "c1") %>%
  kable()

```
### 3.2.2 High School Completion Rate vs. Number of Domestic Incidence

To examine the relationship between high school completion rate and numbers of domestic incidence, scatterplot can be plotted as below. The graph indicates that less high school completion rate is associated with more numbers of domestic incidence per 100,000 people. In other words, people who have not graduated from high school is correlated to more likelyhood of getting involved in domestic incidence. 
```{r High_School_Completion_Rate_vs._Number_of_Domestic_Incidence}
full_data %>%
  select(graduated, fincident_rate) %>%
  ggplot(mapping = aes(x = graduated, y = fincident_rate)) +
  geom_point(size = 2, color = "steelblue", alpha = 0.7) +
  geom_smooth(color = "darkred") +
    labs(title = "Figure 3.2.2: High School Completion Rate vs. Domestic Incidence cases/100,000",
         subtitle = "(2016)",
       x = "High School Completion Rate", 
       y = "Domestic Incidence cases/100,000",
       caption = "Source: Australian bureau of statics (ABS), Victorian Police") +
  theme(plot.title = element_text(size=6)) +
  theme_gray(base_size = 6)
```
**Pearson's r**

Pearson's r is also conducted to examine the correlation. As the result demonstrate, the correlation between the high school completion rate and domestic incidence numbers per 100,000 people is statistically significant *(p<0.01)* and is considered moderate in terms of strength (r = -0.57). In other words, high school completion rate is moderately negatively related to the number of domestic incidence per 100,000 persons. 
```{r}
c_df %>%
  filter(id == "c2") %>%
  kable()

```
**Correlation coefficient matrix**

```{r Correlation_coefficient_matrix}
full_data %>%
  select(low_income, graduated, fincident_rate) %>%
  rename("home Incidence rate" = fincident_rate,
         "HighSchool Graduation rate" = graduated,
         "Low-income Percentage" = low_income
         ) %>%
  ggcorr()
```

# 4 Summary
From the analysis, it has been visually and statistically proven that domestic incidence is more likely to occur in places with higher rate of low-income populations and as well low high school completion rate. The reason might be first, low-income population is more likely to encounter things that may stimulate conflicts, and the resulting negative emotions may lead to a series of violent acts, including provoking domestic violence; second, people with low education generally mean low social skills, lower income and are usually associated with the low-income population, as shown in the Correlation coefficient matrix draw Above, therefore they might be ignorant of the negative outcome associated with domestic violence and hence might more likely to commit one.
The results might be spurious in that, first, the data are separately drew from different database and thus it is skeptical still to directly bind the data together; second, the number of domestic incidence drew from Victoria Police department sliced data collected from June,2015~June,2016 while the AURIN_dataset drew from census data is collected in Aug,2016, thus the data might not be perfectly matching with regards to time. 


