---
title: "Collecting_analyzing_data_web"
author: "Junyou Chen"
date: "3/4/2022"
output: 
  github_document:
    toc: true
    toc_depth: 5
---

```{r setup, include=FALSE}
options(repos = list(CRAN="http://cran.rstudio.com/"))
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

## required packages 
install.packages("countrycode")
library(countrycode)
library(geonames)
library(gapminder)
library(tidyverse)
library(knitr)
library(broom)

## create a new profile to store API key information 
usethis::edit_r_profile(scope = "project")
```

## Part 1
### Assignment Requirement
For the assignment, I want you to replace population with population density and evaluate its relationship with average life expectancy. To do this:

Get the country information using geonames - remember there is a R package for this (see the lecture notes)
Merge gapminder and the country information from geonames
Use left_join from dplyr to merge the tables
Note that you cannot directly do this - gapminder writes the name of countries differently from geonames. To complete the merge, you need a unique key to match observations between the data frames
There is neat little package for R called countrycode that helps solve this problem. countrycode() takes as an input a country’s name in a specific format and outputs it using whatever format you specify.
gapminder stores them using the country.name format
geonames stores them under the countryCode column using the iso2c format
I leave it to you to make the joining operation work
Calculate the population density for each observation
Produce an updated graph using population density
If you want to be real fancy, estimate a statistical model or compare the relationship across continents

### Update the graph (Population Density & LifeExpectancy)
```{r }
key <- getOption("geonamesUsername")

##scrap information from geoname
countryInfo <- GNcountryInfo() %>%
  as_tibble() %>%
  select("countryCode", "areaInSqKm")

##Create a common variable with the iso3c code in each data frame, merge the data, and create a country identifier
countryInfo$iso3c <- countrycode(countryInfo$countryCode, origin = "iso2c", destination = "iso3c")
gapminder$iso3c   <- countrycode(gapminder$country, origin = "country.name", destination = "iso3c")

##Join 2 dfs together using left_join
full_data <- left_join(gapminder,countryInfo, by = "iso3c")

full_data <- full_data %>%
  mutate(areaInSqKm = as.numeric(as.character(areaInSqKm))) %>%
  mutate(density = pop / areaInSqKm)

full_data %>%
  ## use log() function to make the graph more clear 
  ggplot(mapping = aes(x = log(density), y = lifeExp)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  labs(title = "Population Density & Life Expectancy",
    x = "Log Population Density(population per SqKm)",
    y = "Life Expectancy") 

```
As we can see on the graph, generally speaking the higher the population density is, the longer life expectation is. 
We can further examine the correlation between population density and life expectancy using Pearson's r.

### Pearson's R
```{r}
l <- list(
  c1 = cor.test( ~ density + lifeExp, full_data)
)
c_df <- map_dfr(l,tidy,.id = 'id')
c_df %>%
  kable()

```
As the result demonstrate, the correlation between the population density and life expectancy is statistically significant (p<0.01) and is considered weak in terms of strength (r = -0.2). Therefore, we can say that there is a positive correlation betwee n population density and life expancy but the correlation is fairly weak. 

### Correlation Across Continent
To further examine the correlation, we can compare by different continents. 

#### Population Density & LifeExpectancy(Asia)
```{r}

full_data %>%
  filter(continent == "Asia") %>%
  ## use log() function to make the graph more clear 
  ggplot(mapping = aes(x = log(density), y = lifeExp)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  labs(title = "Population Density & Life Expectancy",
       subtitle ="(Asia)",
    x = "Log Population Density(population per SqKm)",
    y = "Life Expectancy") 
```

#### Population Density & LifeExpectancy(Europe)
```{r}
full_data %>%
  filter(continent == "Europe") %>%
  ## use log() function to make the graph more clear 
  ggplot(mapping = aes(x = log(density), y = lifeExp)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  labs(title = "Population Density & Life Expectancy",
       subtitle ="(Europe)",
    x = "Log Population Density(population per SqKm)",
    y = "Life Expectancy") 
```

#### Population Density & LifeExpectancy(Africa)
```{r}

full_data %>%
  filter(continent == "Africa") %>%
  ## use log() function to make the graph more clear 
  ggplot(mapping = aes(x = log(density), y = lifeExp)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  labs(title = "Population Density & Life Expectancy",
       subtitle = "(Africa)",
    x = "Log Population Density(population per SqKm)",
    y = "Life Expectancy") 
```

#### Population Density & LifeExpectancy(Americas)
```{r}

full_data %>%
  filter(continent == "Americas") %>%
  ## use log() function to make the graph more clear 
  ggplot(mapping = aes(x = log(density), y = lifeExp)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  labs(title = "Population Density & Life Expectancy",
       subtitle ="(Americas)",
    x = "Log Population Density(population per SqKm)",
    y = "Life Expectancy") 
```

#### Population Density & LifeExpectancy(Oceania)
```{r}

full_data %>%
  filter(continent == "Oceania") %>%
  ## use log() function to make the graph more clear 
  ggplot(mapping = aes(x = log(density), y = lifeExp)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  labs(title = "Population Density & Life Expectancy",
       subtitle ="(Oceania)",
    x = "Log Population Density(population per SqKm)",
    y = "Life Expectancy") 
```

### Pearson's R (by Continents)

```{r}
data_continent <- full_data %>%
  pivot_wider(names_from = continent, values_from = density) %>%
  mutate_if(is.numeric, ~replace(.,is.na(.), 0)) ##replace NA with 0
l <- list(
  asia = cor.test( ~ Asia + lifeExp, data_continent),
  africa = cor.test( ~ Africa + lifeExp, data_continent),
  europe = cor.test( ~ Europe + lifeExp, data_continent),
  americas = cor.test( ~ Americas + lifeExp, data_continent),
  oceania = cor.test( ~ Oceania + lifeExp, data_continent)
)
continent_df <- map_dfr(l,tidy,.id = 'id')
continent_df %>%
  kable()

```
As we can see from the results of Pearson's r for each continent, all continent examined certain levels of correlation between life expectancy and population density since P-value < 0.01. 
There are 2 things worth noticing however. First, among 5 continents, only Africa examined negative correlation. Second, only Europe examined moderate level of correlation between life expectancy and population density (r = 0.4) while all other four continent exhibit rather weak correlation (-0.12<r<0.15).
