---
title: "Part2_Asteroids_Movie"
author: "Junyou Chen"
date: "3/9/2022"
output: 
  github_document:
    toc: true
    toc_depth: 2
---
## Required Packages
```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(knitr)
library(httr)
library(jsonlite)
library(base)

```

## Asteroids Near Earth
In this part, I tried to practice API query to request information from the website. 
I get data for Asteroids near earth from [Jet Propulsion Laboratory](https://www.jpl.nasa.gov/) by practicing API and unnested the data using `unnest_longer()` and `unnest()`.
[Here](https://ssd-api.jpl.nasa.gov/doc/cad.html) is where I get the API.
The table below only showed the first 6 rows of data. 
```{r}
## retrieve the data

asteroid_api <- function(start, end) {
  # send GET request
  link <- sprintf("https://ssd-api.jpl.nasa.gov/cad.api?date-min=%s&date-max=%s", start, end)
  response <- GET(
    url = link
  )
  # parse response to JSON
  response_df <- content(response)

  return(response_df)
}
## to extract asteroid near earth data from 2020-01-01 to 2021-01-01.
data <- asteroid_api("2020-01-01", "2021-01-01")
## to first store data$data in a dataframe 
data_df <- tibble(data$data)
## to add a new column with repeated values from `fields`
x<-rep(c("des","orbit_id","jd","cd","dist","dist_min","dist_max","v_rel","v_inf","t_sigma_f","h"),times = 1444)

asteroid_df<- data_df %>%
unnest_longer(col=1) %>%
  add_column(x, .before = 1) %>%
  rename(value = `data$data`) %>%
  pivot_wider(names_from = x, values_from = value,values_fn = list) %>%
  unnest(cols = c(des, orbit_id, jd, cd, dist, dist_min, dist_max, v_rel, v_inf, 
    t_sigma_f, h))

kable(head(asteroid_df))
```
Here are further explanations for variables. 
<br/>**des** - primary designation of the asteroid or comet (e.g., 443, 2000 SG344)
<br/>**orbit_id** - orbit ID
<br/>**jd** - time of close-approach (JD Ephemeris Time, TDB)
<br/>**cd** - time of close-approach (formatted calendar date/time, TDB)
<br/>**dist** - nominal approach distance (au)
<br/>**dist_min** - minimum (3-sigma) approach distance (au)
<br/>**dist_max** - maximum (3-sigma) approach distance (au)
<br/>**v_rel** - velocity relative to the approach body at close approach (km/s)
<br/>**v_inf** - velocity relative to a massless body (km/s)
<br/>**t_sigma_f** - 3-sigma uncertainty in the time of close-approach (formatted in days, hours, and minutes; days are not included if zero; example “13:02” is 13 hours 2 minutes; example “2_09:08” is 2 days 9 hours 8 minutes)
<br/>**h** - absolute magnitude H (mag)

